---
title: Full title of the paper (Capitalized)
author:
  - name: Dominik Leutnant
    affil: 1,2,\ddagger,*
    orcid: 0000-0003-3293-2315
  - name: John Doe
    affil: 2, \dagger, \ddagger
affiliation:
  - num: 1
    address: |
      Muenster University of Applied Sciences - 
      Institute for Infrastructure, Water, Resources, Environment
      Correnstr. 25, 48149 Muenster, Germany
    email: leutnant@fh-muenster.de
  - num: 2
    address: |
      Your department
      Street, City, Country
    email: mail@mail.com
# author citation list in chicago format
authorcitation: |
  Leutnant, D.; Doe, J.
# firstnote to eighthnote
firstnote: |
  Current address: Updated affiliation
secondnote: |
  These authors contributed equally to this work.
correspondence: |
  leutnant@fh-muenster.de; Tel.: +XX-000-00-0000.
# document options
journal: notspecified
type: article
status: submit
# front matter
simplesummary: |
  A Simple summary goes here.
abstract: |
  A single paragraph of about 200 words maximum. For research articles, 
  abstracts should give a pertinent overview of the work. We strongly encourage
  authors to use the following style of structured abstracts, but without 
  headings: 1) Background: Place the question addressed in a broad context and
  highlight the purpose of the study; 2) Methods: Describe briefly the main
  methods or treatments applied; 3) Results: Summarize the article's main 
  findings; and 4) Conclusion: Indicate the main conclusions or interpretations. 
  The abstract should be an objective representation of the article, it must not 
  contain results which are not presented and substantiated in the main text and 
  should not exaggerate the main conclusions.
keywords: |
  keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific 
  to the article, yet reasonably common within the subject discipline.).
# back matter
acknowledgement: |
  All sources of funding of the study should be disclosed. Please clearly 
  indicate grants that you have received in support of your research work. 
  Clearly state if you received funds for covering the costs to publish in open 
  access.
authorcontributions: |
  For research articles with several authors, a short paragraph specifying their 
  individual contributions must be provided. The following statements should be 
  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
  limited to those who have contributed substantially to the work reported.
funding: |
  Please add: ``This research received no external funding'' or ``This research 
  was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded 
  by XXX''. Check carefully that the details given are accurate and use the 
  standard spelling of funding agency names at 
  \url{https://search.crossref.org/funding}, any errors may affect your future 
  funding.
institutionalreview: |
  In this section, you should add the Institutional Review Board Statement and 
  approval number, if relevant to your study. You might choose to exclude 
  this statement if the study did not require ethical approval. Please note 
  that the Editorial Office might ask you for further information. Please add 
  “The study was conducted in accordance with the Declaration of Helsinki, 
  and approved by the Institutional Review Board (or Ethics Committee) of 
  NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies 
  involving humans. OR “The animal study protocol was approved by the 
  Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE 
  (protocol code XXX and date of approval).” for studies involving animals. 
  OR “Ethical review and approval were waived for this study due to REASON 
  (please provide a detailed justification).” OR “Not applicable” for
   studies not involving humans or animals.
informedconsent: |
  Any research article describing a study involving humans should contain this 
  statement. Please add ``Informed consent was obtained from all subjects 
  involved in the study.'' OR ``Patient consent was waived due to REASON 
  (please provide a detailed justification).'' OR ``Not applicable'' for 
  studies not involving humans. You might also choose to exclude this statement 
  if the study did not involve humans.
  
  Written informed consent for publication must be obtained from participating 
  patients who can be identified (including by the patients themselves). Please 
  state ``Written informed consent has been obtained from the patient(s) to 
  publish this paper'' if applicable.
dataavailability: |
  We encourage all authors of articles published in MDPI journals to share 
  their research data. In this section, please provide details regarding where 
  data supporting reported results can be found, including links to publicly 
  archived datasets analyzed or generated during the study. Where no new data 
  were created, or where data is unavailable due to privacy or ethical 
  re-strictions, a statement is still required. Suggested Data Availability 
  Statements are available in section “MDPI Research Data Policies” at 
  \url{https://www.mdpi.com/ethics}.
conflictsofinterest: |
  Declare conflicts of interest or state 'The authors declare no conflict of 
  interest.' Authors must identify and declare any personal circumstances or
  interest that may be perceived as inappropriately influencing the
  representation or interpretation of reported research results. Any role of the
  funding sponsors in the design of the study; in the collection, analyses or 
  interpretation of data in the writing of the manuscript, or in the decision to 
  publish the results must be declared in this section. If there is no role, 
  please state 'The founding sponsors had no role in the design of the study; 
  in the collection, analyses, or interpretation of data; in the writing of the 
  manuscript, an in the decision to publish the results'.
supplementary: |
 The following supporting information can be downloaded at:  
 \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.
abbreviations:
  - short: MDPI
    long: Multidisciplinary Digital Publishing Institute
  - short: DOAJ
    long: Directory of open access journals
  - short: TLA
    long: Three letter acronym
  - short: LD 
    long: linear dichroism
bibliography: mybibfile.bib
appendix: appendix.tex
endnotes: false
output: 
  rticles::mdpi_article:
    extra_dependencies: longtable
---

# Version

This Rmd-skeleton uses the mdpi Latex template published 2024-03-21. 
However, the official template gets more frequently updated than the **rticles**
package. Therefore, please make sure prior to paper submission, that you're 
using the most recent .cls, .tex and .bst files 
(available [here](http://www.mdpi.com/authors/latex)).

# Article Header Information

The YAML header includes information needed mainly for formatting the front and 
back matter of the article. Required elements include:

```yaml
title: Title of the paper
author:
  - name: first and last name
    affil: |
      One or more comma seperated numbers corresponding to affilitation
      and one or more  comma seperated symbols corresponding 
      optional notes.
    orcid: optional orcid number
affiliation:  
  - num: 1,..., n for each affiliation
    address: required
    email: required
authorcitation: |
  Lastname, F.
correspondence: |
  email@email.com; Tel.: +XX-000-00-0000.
journal: notspecified
type: article
status: submit
```

Journal options are in Table \ref{tab:mdpinames}. The `status` variable should 
generally not be changed by authors. The `type` variable describes 
the type of of submission and defaults to `article` but can be replaced with any of the ones in Table \ref{tab:mdpitype}

```{r mdpitype, echo = FALSE}
type <- c("abstract, addendum, article, book, bookreview, briefreport, 
casereport, comment, commentary, communication, conferenceproceedings, 
correction, conferencereport, entry, expressionofconcern, extendedabstract, 
datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, 
obituary, opinion, projectreport, reply, retraction, review, perspective, 
protocol, shortnote, studyprotocol, systematicreview, supfile, 
technicalnote, viewpoint, guidelines, registeredreport, tutorial")
knitr::kable(rticles::string_to_table(type, 3), 
             col.names = NULL, format = "latex", booktab = TRUE, 
             caption = "MDPI article types.")
```

## Journal Specific YAML variables

```yaml
# for journal Diversity,
# add the Life Science Identifier using:
lsid: http://zoobank.org/urn:lsid:zoobank.org:act:nnnn


# for journal Applied Sciences
# add featured application
featuredapplication: |
  Authors are encouraged to provide a concise 
  description of the specific application or 
  a potential application of the work. This 
  section is not mandatory.

# for the journal Data
# add dataset doi and license
dataset: https://doi.org/10.1000/182
datasetlicense: CC-BY-4.0

# for the journal Toxins
# add key contributions
keycontributions: |
  The breakthroughs or highlights of the manuscript. 
  Authors can write one or two sentences to describe 
  the most important part of the paper.

# for the journal Encyclopedia
encyclopediadef: |
  For entry manuscripts only: please provide a brief overview
  of the entry title instead of an abstract.
entrylink: The Link to this entry published on the encyclopedia platform.

# for the journal Advances in Respiratory Medicine
# add highlights
addhighlights: |
  This is an obligatory section in “Advances in Respiratory Medicine”, 
  whose goal is to increase the discoverability and readability of the
  article via search engines and other scholars. Highlights should not 
  be a copy of the abstract, but a simple text allowing the reader to 
  quickly and simplified find out what the article is about and what can 
  be cited from it. Each of these parts should be devoted up to 2~bullet 
  points.

```


\startlandscape
```{r, mdpinames, echo=FALSE, results='asis'}
mdpi_journals <- "acoustics, actuators, addictions, admsci, adolescents, aerobiology, aerospace, agriculture, agriengineering, agrochemicals, agronomy, ai, air, algorithms, allergies, alloys, analytica, analytics, anatomia, animals, antibiotics, antibodies, antioxidants, applbiosci, appliedchem, appliedmath, applmech, applmicrobiol, applnano, applsci, aquacj, architecture, arm, arthropoda, arts, asc, asi, astronomy, atmosphere, atoms, audiolres, automation, axioms, bacteria, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomass, biomechanics, biomed, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, blsf, brainsci, breath, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, chips, cimb, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coasts, coatings, colloids, colorants, commodities, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, covid, crops, cryptography, crystals, csmf, ctn, curroncol, cyber, dairy, data, ddc, dentistry, dermato, dermatopathology, designs, devices, diabetology, diagnostics, dietetics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, entomology, entropy, environments, environsciproc, epidemiologia, epigenomes, est, fermentation, fibers, fintech, fire, fishes, fluids, foods, forecasting, forensicsci, forests, foundations, fractalfract, fuels, future, futureinternet, futurepharmacol, futurephys, futuretransp, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, grasses, gucdd, hazardousmatters, healthcare, hearts, hemato, hematolrep, heritage, higheredu, highthroughput, histories, horticulturae, hospitals, humanities, humans, hydrobiology, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijpb, ijtm, ijtpp, ime, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jal, jcdd, jcm, jcp, jcs, jcto, jdb, jeta, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jvd, jzbg, kidneydial, kinasesphosphatases, knowledge, land, languages, laws, life, liquids, literature, livers, logics, logistics, lubricants, lymphatics, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, merits, metabolites, metals, meteorology, methane, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, microplastics, minerals, mining, modelling, molbank, molecules, mps, msf, mti, muscles, nanoenergyadv, nanomanufacturing, nanomaterials, ncrna, ndt, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, nri, nursrep, nutraceuticals, nutrients, obesities, oceans, ohbm, onco, oncopathology, optics, oral, organics, organoids, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacoepidemiology, pharmacy, philosophies, photochem, photonics, phycology, physchem, physics, physiologia, plants, plasma, platforms, pollutants, polymers, polysaccharides, poultry, powders, preprints, proceedings, processes, prosthesis, proteomes, psf, psych, psychiatryint, psychoactives, publications, quantumrep, quaternary, qubs, radiation, reactions, receptors, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, rheumato, risks, robotics, ruminants, safety, sci, scipharm, sclerosis, seeds, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, software, soilsystems, solar, solids, spectroscj, sports, standards, stats, std, stresses, surfaces, surgeries, suschem, sustainability, symmetry, synbio, systems, targets, taxonomy, technologies, telecom, test, textiles, thalassrep, thermo, tomography, tourismhosp, toxics, toxins, transplantology, transportation, traumacare, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, venereology, vetsci, vibration, virtualworlds, viruses, vision, waste, water, wem, wevj, wind, women, world, youth, zoonoticdis"

knitr::kable(rticles::string_to_table(mdpi_journals, 8),
             col.names = NULL,
             format = "latex",
             booktabs = TRUE, 
             caption = "MDPI journal names.",
             longtable = TRUE)

```
\finishlandscape

# Importación
Pongo este apartado aquí no porque esté segura de que es el primero sino para que se vea más, he hecho algunas pruebas para importar uno de los dataframes lo voy dejando aquí por si sirve


```{r}
# ESTO EN TEORÍA FUNCIONA, para leerlos todos
library('stringr')

variables <- c('Nombre', 'Nac.', 'Nac./1000')
ficheros <- paste0('../data/',list.files('../data'))
# Seleccionar solo los archivos que sabemos que tienen los datos de nombres
ficheros <- grep('nombres.*[.]csv',ficheros, value=T)
# Esto recoge la parte numérica al final del fichero que es donde están las últimas dos cifras de la década
decadas <- gsub("[.]csv","",str_extract(list.files('../data'),'[a-z0-9\\-]+[.]csv'))

# Elimina los menores de 30 y convierte los números a década completa
decadas <- sapply(decadas, function(x) {
  # Si es menor30, se descarta devolviendo NA
  if (grepl("menor30", x, ignore.case = TRUE)) {
    return(NA)
  }
  # Si es 30-39 lo pasamos a 1930
  if (grepl("30-39", x)) {
    return("1930")
  }
  # Extrae las dos últimas cifras para convertirlas a década completa
  num <- as.numeric(str_extract(x, "\\d{2}$"))
  if (is.na(num)) return(NA)
  
  # Partimos de 1930 hasta 2020
  if (num >= 30) {
    return(as.character(1900 + num))
  } else {
    return(as.character(2000 + num))
  }
})
# Elimina los NAs correspondientes a menor30
valid_dec <- !is.na(decadas)
ficheros <- ficheros[valid_dec]
decadas <- decadas[valid_dec]


PrimerH <- TRUE
PrimerM <- TRUE
i <- 1
while(i<=length(ficheros)){
  # Lee cada fichero en una variable auxiliar
  auxdf <- read.csv(ficheros[i], header = F, sep='\t', skip=7, fileEncoding = 'Windows-1252', dec=',')
  # Convierte los miles realmente a miles (el . hacía que fuera considerado string)
  auxdf[,'V3'] <- as.numeric(gsub('[.]','',auxdf[,'V3']))
  # Sale una columna solo con el numero de nombre 1,2,3... la quitamos
  auxdf <- auxdf[,-1]
  colnames(auxdf) <- c(c('Nombre'),paste0(variables[c(2,3)],'_',decadas[i]))
  
  # Eliminar las filas que salen con NA (hacer ahora porque si se hace en el grande puede borrar algunas que tengan NA solo   en algunos años)
  auxdf <- auxdf[!(rowSums(is.na(auxdf))>0),]
  # Juntaremos todos los datasets de nombres de hombre y luego todos los de mujer, al final ya lo uniremos todo
  if(grepl('H',ficheros[i])){ # Identifica qué datos son de hombres porque hay H en el nombre del fichero, para mujeres M
    if(PrimerH){
      TotalH <- auxdf
      PrimerH <- FALSE
    } else{
      # El resto se unen a partir de coincidencias en la columna de nombres
      TotalH <- merge(x = TotalH, y = auxdf, by = "Nombre", all=TRUE)
    }
  } else if(grepl('M',ficheros[i])){
    if(PrimerM){
      TotalM <- auxdf
      PrimerM <- FALSE
    } else{
      TotalM <- merge(x = TotalM, y = auxdf, by = "Nombre", all=TRUE)
    }
  }
  
  i <- i+1
}
# Añadimos aquí las columnas de sexo ya que sabemos cuál corresponde a cada grupo de datos
TotalM$Sexo <- 'Mujer'
TotalH$Sexo <- 'Hombre'

# Dataset final total uniendo el de hombres y mujeres (luego vemos como ordenamos filas y columnas)
Nombres <- rbind(TotalH,TotalM)
Nombres$Sexo <- factor(Nombres$Sexo)

# Lo voy a guardar y eliminar las variables que no usamos más porque son grandes, ocuparán bastante
save(Nombres, file = '../data/NombresSucio.RData')
rm(list = ls())
load('../data/NombresSucio.RData')
```

# Convertir df a formato tidy
```{r}
library(tidyr)
library(dplyr)

# Pasamos a longer para poder manejar mejor los datos a la hora de representarlos o hacer cálculos estadísticos
Nombres_longer <- Nombres %>%
  relocate(`Nac._2000`:`Nac./1000_2020`, .after = `Nac./1000_1990`) %>%  # Reordenamos primero las columnas de 1930 a 2020
  pivot_longer(cols = -c(Nombre, Sexo), names_to = "TipoDecada", values_to = "Valor") %>%
  separate(TipoDecada, into = c("Tipo", "Decada"), sep = "_") # Dividimos las dos variables de nacimientos en dos columnas

# Convertimos a wider para poder obtener por separado ambas variables de nacimientos
Nombres_tidy <- Nombres_longer %>% 
  pivot_wider(names_from = "Tipo", values_from = "Valor")
```

# Pruebas de evoluciones temporales
```{r}
# Ranking de nombres por década y sexo
Decada_2000 <- Nombres_tidy %>%
  filter(Decada == 2000, Sexo == "Hombre") %>% 
  select(Nombre, `Nac./1000`, Sexo) %>%
  arrange(desc(`Nac./1000`)) %>% 
  group_by(Sexo)

# Evolución de un nombre a lo largo del tiempo
Nombre_evol <- Nombres_tidy %>%
  filter(Nombre == str_to_upper("Marc")) %>% 
  select(Nombre, Decada, `Nac./1000`, Sexo) %>%
  arrange(Decada) %>% 
  group_by(Decada)

print(Nombre_evol) # Al acabar la dictadura crecen rápidamente nombres valencianos como Marc 


# Función para ver qué nombres aparecen en ambos sexos
Nombres_nb <- function(decada){
  Nombres_tidy %>%
    filter(Decada == decada) %>%
    distinct(Nombre, Sexo) %>% 
    group_by(Nombre) %>% 
    filter(n() == 2) %>% 
    pull(Nombre) %>% 
    unique()
}

# No parece dividirse por décadas
Nombres_nb(1930)
```

# Gráficos tendencias nombres
```{r}
library(ggplot2)

# Función que genera un gráfico a partir de los datos y variables introducidos
grafico_evolucion <- function(data, y_var, y_lab, title, subtitle, y_format = waiver()){
  
  # Convierte el nombre de la columna de forma que aes() lo pueda interpretar 
  y_sym <- sym(y_var)
  
  # Convierte la columna "Década" en numérico para que el eje X sea continuo
  plot_data <- data %>%
    mutate(Decada = as.numeric(Decada))
  
  ggplot(plot_data, aes(x = Decada, y = !!y_sym, color = Sexo, group = Sexo)) + ##!!y_sym usa la columna correspondiente a y_var
    geom_line(linewidth = 1) +
    geom_point(size = 2.5) +
    labs(
      title = title,
      subtitle = subtitle,
      x = "Década",
      y = y_lab,
      color = "Sexo"
    ) +
    # Asegura que todas las décadas (1930, 1940, etc.) aparezcan en el eje X
    scale_x_continuous(breaks = seq(min(plot_data$Decada, na.rm = TRUE), 
                                    max(plot_data$Decada, na.rm = TRUE), 
                                    by = 10)) +
    # Permite un formato personalizado para el eje Y (para usar "%")
    scale_y_continuous(labels = y_format) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5),
    )
}

## Nacimientos por década

# Nacimientos totales agrupando por "Decada" y "Sexo"
nacimientos_por_sexo <- Nombres_tidy %>%
  group_by(Decada, Sexo) %>%
  summarise(TotalNacimientos = sum(Nac., na.rm = TRUE), .groups = "drop")

# Calcula el total de "Ambos" sumando los resultados ya calculados
nacimientos_ambos <- nacimientos_por_sexo %>%
  group_by(Decada) %>%
  summarise(TotalNacimientos = sum(TotalNacimientos, na.rm = TRUE)) %>%
  mutate(Sexo = "Ambos")

# Une los dataframes (por sexo y "Ambos") en uno solo y ordena
Total_Nacimientos_Decada <- bind_rows(nacimientos_por_sexo, nacimientos_ambos) %>%
  arrange(Decada, Sexo)

grafico_evolucion(
  data = Total_Nacimientos_Decada,
  y_var = "TotalNacimientos",
  y_lab = "Número Total de Nacimientos",
  title = "Evolución del Número Total de Nacimientos",
  subtitle = "Por década y género (1930-2020)",
)


## Longitud media de nombres por década

# Filtra filas sin nacimientos y calcula el número de carácteres de cada nombre
Nombres_longitud <- Nombres_tidy %>%
  filter(!is.na(Nac.) & Nac. > 0) %>%
  mutate(Longitud = nchar(Nombre))

# Calcula la media ponderada por sexo. Los nombres más comunes tienen más peso en la media
medias_por_sexo <- Nombres_longitud %>%
  group_by(Decada, Sexo) %>%
  summarise(LongitudMedia = weighted.mean(Longitud, w = Nac.), .groups = "drop")

# Para ambos sexos se recalcula la media
medias_ambos_sexos <- Nombres_longitud %>%
  group_by(Decada) %>%
  summarise(LongitudMedia = weighted.mean(Longitud, w = Nac.), .groups = "drop") %>%
  mutate(Sexo = "Ambos")

Longitud_media_comb <- bind_rows(medias_por_sexo, medias_ambos_sexos) %>%
  arrange(Decada, Sexo)

grafico_evolucion(
  data = Longitud_media_comb,
  y_var = "LongitudMedia",
  y_lab = "Longitud Media del Nombre",
  title = "Evolución de la Longitud Media de los Nombres",
  subtitle = "Ponderada por el número de nacimientos (1930-2020)"
)


## Porcentaje nombres compuestos por década y sexo

# Filtra solo los nombres que contienen un espacio
Nombres_comp <- Nombres_tidy %>%
  filter(grepl(" ", Nombre), !is.na(Nac.) & Nac. > 0)

# Suma los nacimientos de estos nombres compuestos, por sexo
comp_por_sexo <- Nombres_comp %>%
  group_by(Decada, Sexo) %>%
  summarise(TotalNacimientosCompuestos = sum(Nac., na.rm = TRUE), .groups = "drop")

# Suma los totales por sexo para obtener el total de "Ambos"
comp_ambos <- comp_por_sexo %>%
  group_by(Decada) %>%
  summarise(TotalNacimientosCompuestos = sum(TotalNacimientosCompuestos, na.rm = TRUE)) %>%
  mutate(Sexo = "Ambos")

Nacimientos_comp <- bind_rows(comp_por_sexo, comp_ambos) %>%
  arrange(Decada, Sexo)

# Calcular el porcentaje
Porcentaje_Comp_Decada <- Total_Nacimientos_Decada %>%
  left_join(Nacimientos_comp, by = c("Decada", "Sexo")) %>%
  mutate(
    TotalNacimientosCompuestos = replace_na(TotalNacimientosCompuestos, 0)
  ) %>%
  mutate(
    PorcentajeCompuestos = (TotalNacimientosCompuestos / TotalNacimientos) * 100
  ) %>%
  select(Decada, Sexo, PorcentajeCompuestos)

grafico_evolucion(
  data = Porcentaje_Comp_Decada,
  y_var = "PorcentajeCompuestos",
  y_lab = "Porcentaje de Nacimientos (%)",
  title = "Evolución del Porcentaje de Nombres Compuestos",
  subtitle = "Respecto al total de nacimientos de cada grupo (1930-2020)",
  y_format = scales::label_number(accuracy = 0.1, suffix = " %") # Este formato muestra el eje Y como "XX.X %"
)


## Cantidad de nombres distintos por década y sexo

# Filtra filas sin nacimientos
Nombres_variedad <- Nombres_tidy %>%
  filter(!is.na(Nac.) & Nac. > 0)

# Calcula la cantidad de nombres distintos por década y sexo
variedad_sexo <- Nombres_variedad %>%
  group_by(Decada, Sexo) %>%
  summarise(NombresDistintos = n_distinct(Nombre), .groups = "drop")

variedad_ambos <- Nombres_variedad %>%
  group_by(Decada) %>%
  summarise(NombresDistintos = n_distinct(Nombre), .groups = "drop") %>% 
  mutate(Sexo = "Ambos")

Variedad_Nombres_Decada <- bind_rows(variedad_sexo, variedad_ambos) %>%
  arrange(Decada, Sexo)

grafico_evolucion(
  data = Variedad_Nombres_Decada,
  y_var = "NombresDistintos",
  y_lab = "Nombres Distintos",
  title = "Evolución de la Variedad de Nombres Distintos",
  subtitle = "Por década y género (1930-2020)"
)


## Nuevos nombres por década y sexo

Nombres_nuevos_decada <- Nombres_tidy %>%
  filter(!is.na(Nac.) & Nac. > 0)

Nombres_unicos_decada <- Nombres_nuevos_decada %>%
  group_by(Decada, Sexo, Nombre) %>%
  mutate(Decada = as.numeric(Decada))

Nombres_primera_aparicion <- Nombres_unicos_decada %>%
  group_by(Nombre, Sexo) %>%
  summarise(PrimeraDecada = min(Decada, na.rm = T), .groups = "drop")

nuevos_nombres_decada <- Nombres_primera_aparicion %>% 
  group_by(Decada = as.character(PrimeraDecada), Sexo) %>% 
  summarise(NuevosNombres = n(), .groups = "drop")

nuevos_nombres_ambos <- nuevos_nombres_decada %>% 
  group_by(Decada) %>% 
  summarise(NuevosNombres = sum(NuevosNombres, na.rm = T), .groups = "drop") %>% 
  mutate(Sexo = "Ambos")

Nombres_Nuevos <- bind_rows(nuevos_nombres_decada, nuevos_nombres_ambos) %>%
  arrange(Decada, Sexo)

grafico_evolucion(
  data = Nombres_Nuevos,
  y_var = "NuevosNombres",
  y_lab = "Nombres Nuevos",
  title = "Cantidad de Nombres Nuevos",
  subtitle = "Por década y género (1930-2020)"
)


## Número de nombres desaparecidos con el tiempo

# Utilizamos el df del caso anterior para ver la última aparición de cada nombre
Nombres_ultima_aparicion <- Nombres_unicos_decada %>%
  group_by(Nombre, Sexo) %>%
  summarise(UltimaDecada = max(Decada, na.rm = TRUE), .groups = "drop")

# Excluimos 2020 porque es la última década que tenemos disponible
max_decada_total <- max(Nombres_unicos_decada$Decada, na.rm = TRUE)

Nombres_desaparecidos <- Nombres_ultima_aparicion %>%
  filter(UltimaDecada < max_decada_total)

desaparecidos_por_sexo <- Nombres_desaparecidos %>%
  group_by(Decada = as.character(UltimaDecada), Sexo) %>%
  summarise(NombresDesaparecidos = n(), .groups = "drop")

desaparecidos_ambos <- desaparecidos_por_sexo %>%
  group_by(Decada) %>%
  summarise(NombresDesaparecidos = sum(NombresDesaparecidos, na.rm = TRUE), .groups = 'drop') %>%
  mutate(Sexo = "Ambos")

Total_Nombres_Desaparecidos <- bind_rows(desaparecidos_por_sexo, desaparecidos_ambos) %>%
  arrange(Decada, Sexo)

grafico_evolucion(
  data = Total_Nombres_Desaparecidos,
  y_var = "NombresDesaparecidos",
  y_lab = "Número de Nombres Desaparecidos",
  title = "Nombres que Desaparecen por Década",
  subtitle = "Nombres vistos por última vez en esa década (y que no reaparecen)",
)

```

```{r}
#Ver los 5 nombres más populares de mujeres en cada año 
Nom_populares <- Nombres_tidy %>%
  group_by(Sexo, Decada) %>%
  arrange(desc(Nac.)) %>% #ordena descendentemente (cantidad de mujeres nacidas con ese nombre)
  slice(1:5) %>%
  ungroup()
print(Nom_populares)

#Gráficos para ver los 10 nombres más comunes por año (5 mujeres y 5 hombres)
df_1930 <- Nom_populares %>%
  filter(Decada == '1930')
g1930<-ggplot(data=df_1930, aes(x=reorder(Nombre, Nac.), y=Nac.)) + geom_col(aes(fill = Sexo)) + facet_grid(~ Sexo, scales = "free_x") + xlab('Nombres') + ylab('Número de Nacimientos') + ggtitle('5 Nombres más populares en 1930 en la provincia de Valencia') + theme(axis.text.x = element_text(angle=45, hjust = 1))
g1930
```


# Introduction

The introduction should briefly place the study in a broad context and highlight
why it is important. It should define the purpose of the work and its
significance. The current state of the research field should be reviewed
carefully and key publications cited. Please highlight controversial and
diverging hypotheses when necessary. Finally, briefly mention the main aim of
the work and highlight the principal conclusions. As far as possible, please
keep the introduction comprehensible to scientists outside your particular
field of research. Citing a journal paper [@bertrand-krajewski_distribution_1998;
@leutnant_stormwater_2016]. And now citing a book reference @gujer_systems_2008.
Some MDPI journals use Chicago and others use APA, this template should choose
the correct citation format for you once you specify the journal in the YAML
header.

To use endnotes, change `endnotes: true` in the YAML header, then use
`\endnote{This is an endnote.}`.


# Materials and Methods

Materials and Methods should be described with sufficient details to allow 
others to replicate and build on published results. Please note that publication 
of your manuscript implicates that you must make all materials, data, computer 
code, and protocols associated with the publication available to readers. Please
disclose at the submission stage any restrictions on the availability of 
materials or information. New methods and protocols should be described in 
detail while well-established methods can be briefly described and appropriately
cited.

Research manuscripts reporting large datasets that are deposited in a publicly 
available database should specify where the data have been deposited and provide
the relevant accession numbers. If the accession numbers have not yet been
obtained at the time of submission, please state that they will be provided
during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require
ethical approval must list the authority that provided approval and the
corresponding ethical approval code.

# Results

This section may be divided by subheadings. It should provide a concise and
precise description of the experimental results, their interpretation as well
as the experimental conclusions that can be drawn.

## Subsection Heading Here

Subsection text here.

### Subsubsection Heading Here

Bulleted lists look like this:

* First bullet
* Second bullet
* Third bullet

Numbered lists can be added as follows:

1. First item
2. Second item
3. Third item

The text continues here.

## Figures, Tables and Schemes

All figures and tables should be cited in the main text as Figure \ref{fig:fig1}, 
\ref{tab:tab1}, etc. To get cross-reference to figure generated by R chunks 
include the `\\label{}` tag in the `fig.cap` attribute of the R chunk: 
`fig.cap = "Fancy Caption\\label{fig:plot}"`.

```{r fig1, echo=FALSE, fig.width=3, fig.cap="A figure added with a code chunk.\\label{fig:fig1}"}
x = rnorm(10)
y = rnorm(10)
plot(x, y)
```


When making tables using `kable`, it is suggested to use
the `format="latex"` and `tabl.envir="table"` arguments
to ensure table numbering and compatibility with the mdpi
document class.

```{r tab1, echo=FALSE}
knitr::kable(mtcars[1:5, 1:3], format = "latex", 
             booktabs = TRUE, 
             caption = "This is a table caption. Tables should be placed in the 
             main text near to the first time they are~cited.", 
             align = 'ccc', centering = FALSE,
             table.envir = "table", position = "H")
```



For a very wide table, landscape layouts are allowed.


\startlandscape

```{r tab2, echo=FALSE}
df <- data.frame(`Title 1` = c("Entry 1", "Entry 2"),
                 `Title 2` = c("Data", "Data"),
                 `Title 3` = c("Data", "Data"),
                 `Title 4` = c("This cell has some longer content that runs over
                               two lines",
                               "Data"))
knitr::kable(df, format = "latex", 
             booktabs = TRUE, 
             caption = "This is a very wide table", 
             align = 'ccc', centering = FALSE,
             table.envir = "table", position = "H")
```

\finishlandscape

## Formatting of Mathematical Components

This is an example of an equation:

$$
a = 1.
$$

If you want numbered equations use Latex and wrap in the equation environment:

\begin{equation}
a = 1,
\end{equation}

the text following an equation need not be a new paragraph. Please punctuate 
equations as regular text.

```{comment}
If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
```

This is the example 2 of equation:

\begin{adjustwidth}{-\extralength}{0cm}
\begin{equation}
a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + 
u + v + w + x + y + z
\end{equation}
\end{adjustwidth}

Theorem-type environments (including propositions, lemmas, corollaries etc.) 
can be formatted as follows:

Example of a theorem:

::: {.Theorem latex=true}
Example text of a theorem
:::

The text continues here. Proofs must be formatted as follows:

Example of a proof:

::: {.proof latex="[Proof of Theorem1]"}
Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is 
clear which theorem is being referred to.
:::

The text continues here.

# Discussion

Authors should discuss the results and how they can be interpreted in 
perspective of previous studies and of the working hypotheses. The findings and 
their implications should be discussed in the broadest context possible. Future 
research directions may also be highlighted.

# Conclusion

This section is not mandatory, but can be added to the manuscript if the
discussion is unusually long or complex.

# Patents

This section is not mandatory, but may be added if there are patents resulting
from the work reported in this manuscript.
